Sketch recogniton and understanding

### 2009-02-18 trunk/draft/ChapterIntroduction.tex
now i am reviewing

%\usepackage{graphicx}
\chapter {Introduciton}
\pagenumbering{arabic}

Computer research nowadays aimed at facilitating computer human interaction in every way. Pen-based interface gives the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces. However, until now there are hardly any complete pen based computer systems. Generally, Current Pen-Based systems make use of the Pen or stylus to perform the same role as the mouse / keyboard. The minority of such systems are equipped with handwriting or digit recognition that is typically restricted for specific language.

 Despite the fact that people frequently express their ideas through drawing them there isn't a system that can fully understand those sketches. Lately, only a few researchers were concerned with the problem although it is highly important.  
Firstly, we need to know;
 \textit { What are the main goals of Sketch understanding systems?}

The main goal of sketch understating systems is to build a system that will interpret the sketches the user draw. The systems ought to be natural as possible to provide user with paper and pencil impression. Furthermore, interpreted sketches are expected to transform into objects, which can me manipulated, modeled and/or simulated.

Sketch here stand for any hand drawn drawing the user draws using digital stylus or similar device. Normally, sketch are composed of smaller segments we call strokes. The time between pen-up and pen-down events identify each stroke.  The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.

Sketch recognition is procedure that will input a sequence of points from user then segment them into clusters of strokes and recognize those clusters to primitive geometric objects and shapes. However, Sketch understanding is transforming the sketch from a group of primitive shapes into comprehendible identified symbols consistent with each other.
% the entire sketched symbols.

@@

People normally sketch thoughts rather than expressing them linguistically. For instance engineers mostly drafts their early designs using paper and pencil before using CAD tools.  People in general and engineers specially make use of various sketches during design process as instance organizational charts, informal drafting and formal design sketches \cite{}. The sketching process itself go thorough numerous stages. Firstly, early stages with rough drafting designs or notations which are mainly done using paper and pencil. After those primary rough sketches are edited and altered until they are stabilized and finally they are transformed into the computer.

 Paper and Pencil are preferred in design as they offer a remarkable ease in editing and altering the primary rough sketches. Current CAD systems don't give the designer the ease of altering the early designs. Designer had to sketch primary designs on paper and pencil next, the finally design is transferred into computer using a menu CAD system. The process is time consuming and tiring. The need for fast design and development process is driven by the fast growing industries and technology.

@@

\section{Problems and challenges}
\label{sec:ProblemsAndChallenges}

@@

\subsection{Search space}
\label{sec:Searchspace}

The user drawn symbols can be having more than one interpolation. In addition, those interpolations increase exponentially with the number of strokes the user draw and as the sketch expands the number of interpolation escalate. This intensifies the search space that the program search to resolve the ambiguities and find best interpolation.

\subsection{Real time}
\label{sec:Realtime}

There are mainly three types of sketch understanding system offline \cite{Vibratory8}, online and interactive. Both online and interactive recognition like \cite{incrmentintention41, SmartSketch56} need the processing after each stroke to be real time so the user does not feel the processing.

>>>>

%\usepackage{graphicx}
\chapter {Introduction}
\pagenumbering{arabic}

Computer research nowadays aimed at facilitating computer human interaction in every way. Pen-based interfaces give the user a pencil-paper like feeling that enhances interaction more than the current used keyboard-mouse computer interfaces. However, until now there are hardly any complete pen based computer systems. Generally, Current Pen-Based systems make use of the Pen or stylus to perform the same role as the mouse / keyboard. A minority of such systems are equipped with handwriting or digit recognition that is typically restricted for specific language.

\section{Sketch Recognition}
Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated.

Sketchs here stand for any hand drawn drawing the user draws using digital stylus or similar device. Normally, sketch are composed of smaller segments we call strokes. The time between pen-up and pen-down events identify each stroke.  The strokes are mainly the path of points, which the user path through between the pen-down and pen-up events while using the stylus or similar device.

%Sketch recognition is procedure that will input a sequence of points from user then segment them into clusters of strokes and recognize those clusters to primitive geometric objects and shapes. However, Sketch understanding is transforming the sketch from a group of primitive shapes into comprehendible identified symbols consistent with each other.
% the entire sketched symbols.
%Sketch recognition is defined as the process of identifying symbols that users draw using single or multiple strokes. Users draw strokes using a pen and the system should immediately interprets their strokes into objects that can be easily manipulated. This paper uses Particle Swarm Algorithm (PSO) to divide the strokes the user draws into meaningful geometric primitives. These geometric primitives are grouped to formulate symbols which are further identified. The results show that using PSO improves segmentation results which guide the symbol recognition phase. This paper uses Support Vector Machine(SVM) classifier which further improves the final recognition accuracy.  %u Final recognition has improved accuracy after the PSO algorithm using a SVM classifier.  
 %Hand drawn sketches are the natural method of exchanging thoughts between engineers and humans in general. Nevertheless, researchers barely give attention to improve current gesture and sketch recognition systems. In a sketch understanding system, users need to feel the freedom of paper and pencil along with the flexibility of computer manipulation.

 %A picture speak a thousand words is also true for sketches. 
%Scientists generally and engineers specifically express thoughts and designs using sketches. Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking. Engineers use paper and pencil design in early stages of design. Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers. Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}. Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.

%Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding. The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation. In segmentation phase, the system divides the stroke into a set of simple geometrical primitives or segments. The third phase, sketch recognition, is clustering strokes and segments to formulate symbols that can be recognized by a classifier system. Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols. 
%the basic symbols are identified in consistent with the whole sketch. \\
%from the user information from the users converts the strokes the user draw into segments which is recognized into basic symbols that can be easily manipulated. Finally, the basic symbols are interpreted in consistent with the whole sketch. \\
%This paper introduces a sketch recognition system. The presented system uses' particle swarm algorithm (PSO) to optimally segment users strokes. Users can draw symbols using any number of strokes in any order they like. The system segments stroke user draws then cluster symbols together then passes them to the SVM symbol recognizer. The experiments in section \ref{sec:Experiments} shows that the proposed system improves the final symbol recognition accuracy.

%The remaining of this papers is organized as follows; section \ref{Sec:RelatedWork} provides a brief literature review.  Section \ref{PSO} explains the general particle swarm algorithm. The proposed system block diagram along with a detailed system description is presented in Section \ref{Sysdisc}. The experiments and results achieved are explained in Section \ref{sec:Experiments}. Finally, section \ref{ConclusionandFutureWork} states the conclusion and future work. 
%The next section will relate previous work in the field. The remaining of the paper will be as following; Section \ref{PSO} will explain the general particle swarm algorithm which is used in the presented system. Section \ref{Sysdisc} will describe the system block diagram and then section \ref{Prepross} will explain the preprocessing step, section \ref{seg} will describe the segmentation algorithm in details. Finally, section \ref{sec:Experiments} will describe the experiments that was preformed.\\

@@

Scientists generally and engineers specifically express thoughts and designs using sketches. Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking. Engineers use paper and pencil design in early stages of design. The sketching process itself go thorough numerous stages. Firstly, early stages with rough drafting designs or notations which are mainly done using paper and pencil. After those primary rough sketches are edited and altered until they are stabilized and finally they are transformed into the computer.

%Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers. Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}. Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.

%People normally sketch thoughts rather than expressing them linguistically. For instance engineers mostly drafts their early designs using paper and pencil before using CAD tools.  People in general and engineers specially make use of various sketches during design process as instance organizational charts, informal drafting and formal design sketches \cite{}. The sketching process itself go thorough numerous stages. Firstly, early stages with rough drafting designs or notations which are mainly done using paper and pencil. After those primary rough sketches are edited and altered until they are stabilized and finally they are transformed into the computer.

 Paper and Pencil are preferred in design as they offer a remarkable ease in editing and altering the primary rough sketches. Current CAD systems do not give the designer the ease of altering the early designs. Designer had to sketch primary designs on paper and pencil next, the finally design is transferred into computer using a menu CAD system. The process is time consuming and tiring. The need for fast design and development process is driven by the fast growing industries and technology.

@@

\section{Applicability }
% talk about application like cad systems , software engineering uml diagrams ....
% meeting idea brainstorming , mind mapping string ..
% communication program  
% sketching programs ...
Application of sketch recognition are ranging from engineering device simulation and design. Drawing and sketching programs are most likely to benefit from integrating a sketch recognition system into there design environments. Other applications include object retrieval, map navigations and various engineering design systems.

\section{Research Scope}
%rewirte this part.............................
%  research will invistgated till sketch recognition in domain of uml diagrams. 
%This research will include investigating the best method to recognize sketches.  The scope of the research will include recognizing various sketches. The methods used in this research are comparing current systems and enhancing them and adobpiting them into simple geometrical symbols.  Using well established AI methods specially \textit{Swarm Intelligence} to solve the problem.

\section{Problem Definition}
\label{sec:ProblemsAndChallenges}

%Sketch recognition is divided into three main steps preprocessing, segmentation and symbol recognition. The preprocessing phase captures user input stroke points and collects basic information about the stroke then proceeds to remove noise and compute basic statistical and geometrical information. In segmentation phase, the strokes are divided into a set of simple geometrical primitives or segments. In the third phase, sketch recognition, strokes and segments are clustered to formulate symbols that can be recognized by a classifier system.

@@

\section{Thesis Overview}
This thesis is arranged such that chapter \ref{} presents a survey of different components of the
sketch recognition. In chapter \ref{}, more light is shed on the feature filter technique for dimensionality
reduction. This is in conjunction with a detailed description of the proposed methods to
enhance the feature filtering. Comparative studies among the proposed methods and stateof-
arts methods are presented in chapter 4. Finally, chapter 5 presents the conclusions of
this thesis and suggested future work.
%\subsubsection{Search space}
%\label{sec:Searchspace}
%
%
%The user drawn symbols can be having more than one interpolation. In addition, those interpolations increase exponentially with the number of strokes the user draw and as the sketch expands the number of interpolation escalate. This intensifies the search space that the program search to resolve the ambiguities and find best interpolation.
%
%\sublsubsection{Real time}
%\label{sec:Realtime}
%
%
%There are mainly three types of sketch understanding system offline \cite{Vibratory8}, online and interactive. Both online and interactive recognition like \cite{incrmentintention41, SmartSketch56} need the processing after each stroke to be real time so the user does not feel the processing.

%A picture speak a thousand words is also true for sketches. 
%Scientists generally and engineers specifically express thoughts and designs using sketches. Engineers use sketches to exchange designs as a natural method of communication rather than writing or speaking. Engineers use paper and pencil design in early stages of design. Design engineers need a powerful computer based system with symbol recognition to help them design, manipulate and store sketches more effectively than using only papers. Increasing the interaction between computers and users in sketch and CAD systems has been the reason for the emerging of few advanced sketch recognition systems \cite{CAD}. Sketch recognition system converts the drawings and informal graphs user draws into objects and symbols that can be used in further interaction with the system.

%Sketch recognition is divided into four main steps preprocessing, segmentation, symbol recognition and finally sketch understanding. The preprocessing process saves user input stroke points from the hardware and collects basic information about the stroke then proceeds to remove noise and primary data calculation. In segmentation phase, the system divides the stroke into a set of simple geometrical primitives or segments. The third phase, sketch recognition, is clustering strokes and segments to formulate symbols that can be recognized by a classifier system. Finally, the sketch understanding phase uses priori information about the sketch to support identification of symbols by excluding any possible invalid symbols. 
%the basic symbols are identified in consistent with the whole sketch. \\
%from the user information from the users converts the strokes the user draw into segments which is recognized into basic symbols that can be easily manipulated. Finally, the basic symbols are interpreted in consistent with the whole sketch. \\

### 2009-02-18 trunk/draft/Results.tex
now i am reviewing

\section{Experiments Setup}
\label{sec:ExperimentsSetup}
\section{Datasets}
\label{sec:Datasets}
\section{Evaluations techniques}
\label{sec:EvaluationsTechniques}

\section{Segmentation Algoirthms Results}
\label{sec:SegmentationAlgoirthmsResults}

\subsection{Swarm Algorithms}
\label{sec:Swarm Algorithms}

\section{Clustering Algorithm Results}
\label{sec:ClusteringAlgorithmResults}

\section{Conclusion}
\label{sec:Conclusion}

\section{ BenchMark Results }
\label{sec:BenchMarkResults}

>>>>

This chapter discribe and illustrate the results generated from the system. In the chapter, section \ref{sec:ExperimentsSetup} will describe the experiments performed on each block of the system. section \ref{sec:Datasets} will describe the datasets used. Section \ref{sec:ResultsDetails} will demonstrate and analysis the result generated.

\section{Experiments Setup}
\label{sec:ExperimentsSetup}
The system was tested using various experiments. The preprocessing procedures and computation was tested by its effects on the segmentation algorithm and was reported in section (\ref{sec:ResultsDetails}). The variable used to tune the swarm algorithm were tested to choose the best parameters to use(see section \ref{sec:PSO}). Segmentation algorithms test was reported and noted in section \ref{sec:SegmentationAlgorithms}.

\section{Datasets}
\label{sec:Datasets}
For classifier and algorithm testing the system used a data set collected by \citeauthor{HeloiseBeautification}. The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
%We divided the dataset into four sets each set contains four different users. Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged.
\begin{figure}[]\centering
\fbox{  \parbox{8cm}{% 
		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
		}
	\caption{The Symbol Set}
	\label{fig:symbolSet}
\end{figure}

%\section{Evaluations techniques}
%\label{sec:EvaluationsTechniques}
%\section {Curvature estimation results}
%\label{sec:Curvatureestimationresults}

\section{Results}
\label{sec:ResultsDetails}

\subsection{PSO Algorithm}
\label{sec:PSO}
The fig. \ref{fig:dataerrorvsiteration.jpg} display the effect of the size of swarm population on the number of vertex reported while segmenting the stroke.\footnote{The fewer the vertex the better the segmentation} The graphs shows that as the number of iteration increase there is a decrease in error until the error reach a saturation and any increase in number of iteration dose not affect the error calculated. Similar behavior is noticed in fig. \ref{fig:datapopulation.jpg}.  These curves and test result in choosing the swarm parameters $c_1,c_2$ and maximum number of iteration and the swarm population. The final values are for maximum number of iteration = 150 with 15 particle for better compensation in the time domain.

\begin{figure}
	\centering
		\includegraphics[scale=0.8]{images/dataerrorvsiteration.jpg.eps}
	\caption{Error vs. iterations }
	\label{fig:dataerrorvsiteration.jpg}
\end{figure}
\begin{figure}
	\centering
		\includegraphics[scale=0.8]{images/datapopulation.jpg.eps}
	\caption{Swarm Population }
	\label{fig:datapopulation.jpg}
\end{figure}

\subsection{Segmentation Algorithms}
\label{sec:SegmentationAlgorithms}
The result of the segmentation algorithm can be viewed in the figs. \ref{fig:results1.JPG}. These result shows the originals stroke and the segmentation that the system generated for the stroke. 
\begin{figure}
	\centering
		\includegraphics{images/results1.JPG.eps}
	\caption{Results Segmentations}
	\label{fig:results1.JPG}
\end{figure}
\begin{figure}
	\centering
		\includegraphics{images/results2.JPG.eps}
	\caption{Results Segmentaion }
	\label{fig:results2.JPG}
\end{figure}
\begin{figure}
	\centering
		\includegraphics{images/results3.JPG.eps}
	\caption{Results }
	\label{fig:results3.JPG}
\end{figure}
\begin{figure}
	\centering
		\includegraphics{images/results4.JPG.eps}
	\caption{Resluts of the segmentations}
	\label{fig:results4.JPG}
\end{figure}

The experiment we implemented was to test the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} shows what the accuracy of each symbol, it clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. Figure \ref{fig:test2} also shows that algorithm 1 alone achieve better performance than algorithm 2 in the symbols that consist of lines only. The combining of both algorithm improves the recognition rate of other symbols.  \\
\begin{figure*}[]
	\centering
		\includegraphics[scale=0.5]{images/test2.eps}
	\caption{The results of second experiment}
	\label{fig:test2}
\end{figure*}
 Another experiment we performed was to test the efficiency svm classifier. We implemented the linear discriminator classifier described in \cite{gestureexample12}(see figure \ref{fig:test3}).  %The result shows that SVM is more scalable than the linear discriminator for this problem. %The test focus on the size of data needed to test each classifier and the achieved accuracy.  %For this experiment we also shows the classification time for both classifiers 
\begin{figure}[]
	\centering
		\includegraphics[scale=0.4]{images/test1.eps}
	\caption{The results of third experiment}
	\label{fig:test3}
\end{figure}
 %Details description of results got from comparing various algorithms  
 %Describe data set used for experiments ( No of data set, No. of categories, No of samples).
 %Describe other Algorithms [Alg 3 in paper \cite{earlyprocess}]used for comparing with the one described in the paper.

 %In the figure \ref{fig:test1}, the accuracy means that final hit rate of all symbols used in the test set.by the  four datasets . 
We performed three experiments to test the system firstly we tested recognition accuracy of shapes in the data set with both algorithms. We also implemented the segmentation algorithm described in \cite{earlyprocess} to use as reference to our swarm algorithms.  As you see in the fig. \ref{fig:test1} shows the accuracy achieved by each algorithm. The results shows that both PSO algorithm achieve better result than other algorithms.  The swarm algorithms were tested with and without the ellipse detection module. The ellipse detection module appear to be superior to results with only the PSO algorithms.  \\  
\begin{figure}[]
	\centering
		\includegraphics[scale=0.3]{images/test1.eps}
	\caption{The results of first experiment}
	\label{fig:test1}
\end{figure}

%\subsection{Clustering Algorithm Results}
%\label{sec:ClusteringAlgorithmResults}

\subsection {Recognition Algorithms }
\label{sec:recognitionAlgorithms}
To test the recognition different set of features are used to select the features that will define symbols better to finally achieve better recognition rate. We tested the system using different types of features. The features explained in section \ref{sec:FeatureExtraction} was divided into different sets; statistical, structural, Rubine features and zernik moments along with composite features. The fig. \ref{fig:testFeatures} shows the result of each set of features when used with the SVM classifier. The results shows that the Rubine features although may achieve good results in single stroke symbols, have a bad performance when used with multi stroke symbols. 
\begin{figure}[]
	\centering
		\includegraphics[scale=0.3]{images/test4.eps}
	\caption{The results of Features experiment}
	\label{fig:testFeatures}
\end{figure}

We test the system using different classifier to see what the effect of the classification step on the final recognition. We also wanted to test the train and test time in both classifiers. We used SVM (RBF) classifiers and the Gaussian Linear classifier used in \cite{}. The fig. \ref{} shows the result of both classifiers with the main three segmentation algorithms. The results shows that the SVM classifier achieve better results. To measure the scalability of the system\footnote{What happens when adding a new symbols to the set of known symbols.} the classifiers are test using different number of symbols (fig. \ref{fig:sampleperCat}). Figures \ref{fig:sampleperCat} demonstrate the scalability of both classifiers by demonstrating the recognition rate in both classifiers with different number of symbols. Figures \ref{fig:sampleperCat},\ref{fig:sampleperCat} shows the time and number of symbols needed for training versus each classifier.

\begin{figure}
	\centering
		\includegraphics{images/sampleperCat.eps}
	\caption{Samples per Category }
	\label{fig:sampleperCat}
\end{figure}

\begin{figure}
	\centering
		\includegraphics[scale=0.7]{images/test4.eps}
	\caption{Tests}
	\label{fig:test4}
\end{figure}

\section{Summary}
\label{sec:ResultSummary}

The result illustrated in this chapter shows the work done on the system and experiments performed. The results shows that the ellipse detection and PSO swarm algorithms gives better performance than other known systems.

%\section{ BenchMark Results }
%\label{sec:BenchMarkResults}
